---
title: "ggmail + forecast = how many emails I will get tomorrow?"
author: "Przemyslaw Biecek"
date: "22/11/2016"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Below you will find information how to download emails from gmail and how to use R to process these emails

Find more details here: http://smarterpoland.pl/index.php/2016/10/ggmail-ggplot2-data-from-gmail-forecast/

## 1. Get emails from gmail

Follow [this link](https://support.google.com/accounts/answer/3024190?hl=en) to download the data from your gmail account as a single mbox file.

It may be large (15GB in my case).

## 2. Filtler out headers

Since I am not going to analyse the content I will remove everything except Subject, Date and Recipent.

The fastest way to do this is to use grep function in the shell

```
cat GGmail.mbox  | grep -e "^Date:" -e "^Subject:" -e "^From:" -e "^To:" > GGheaders.txt
```

## 3. Read the data

Reading and cleaning. Here I am using only timestamps.

```{r, warning=FALSE, message=FALSE}
lines <- readLines("GGheaders.txt")
lines <- grep(lines, pattern="^Date:", value = TRUE)
head(lines)

dates <- substr(lines, 11, 1000)
head(dates)

library("lubridate")
library("forecast")
library("ggplot2")
library("forcats")

daty_clean <- dmy_hms(dates)
head(daty_clean)

# daily agregates
tt <- table(substr(daty_clean, 1, 10))
head(tt)

# dirty dirty dirty
tt <- tt[45:3350]

```

## 4. ggplot it

```{r}
# time series object
tts <- ts(tt, frequency = 356, start=c(2007, 52))

autoplot(tts) +
  geom_smooth(method="loess", span=.03, se=F) +
  xlab("") + ylab("Number of emails per day") +
  ggtitle("Mailbox statistics")

# Some boxplots
tts <- ts(c(tail(tts, 356)), frequency = 356, start=c(2007, 52))
dftts <- data.frame(mails = tts, date=ymd(names(tts)))
dftts$dow <- wday(dftts$date, label = T, abbr = F)
ggplot(dftts, aes(factor(dow), mails, fill=dow)) + geom_boxplot(coef=100) +
  geom_violin(alpha=0.5, scale = "width") +
  theme(legend.position="none") + xlab("") + ylab("Number of mails per day") +
  ggtitle("Monday's flood - mailbox statistics year 2016\n") +
  scale_y_continuous(limits=c(0,260), expand = c(0,0.5)) + coord_flip()


dftts$m <- month(dftts$date, label = T, abbr = F)
dftts$m <- fct_rev(dftts$m)
ggplot(dftts[substr(dftts$date, 1, 4) == "2016",], aes(factor(m), mails, fill=m)) + geom_boxplot(coef=100) +
  geom_violin(alpha=0.5, scale = "width") +
  theme(legend.position="none") + xlab("") + ylab("Number of emails per day") +
  ggtitle("The first month of each semester could be hard \n mailbox statistics year 2016") +
  scale_y_continuous(limits=c(0,260), expand = c(0,0.5)) + coord_flip()

```
